PCA on Finnish Municipalities

### 2022-04-10 MSA-Project.Rmd
Updated variable description table and added references for tables and figures

---
title: "PCA on Finnish Municipalities"
author: "Niko Miller"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: 
  pdf_document:
    keep_tex: false
    number_sections: true
    fig_caption: true
    citation_package: natbib
bibliography: references.bib
header-includes:
- \pagenumbering{gobble}
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsfonts}
- \usepackage{mathtools}
- \usepackage{graphicx}
- \usepackage{placeins}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{framed}
- \usepackage[font={small,it}]{caption}
- \usepackage[font={small,it}]{subcaption}
- \usepackage{hyperref}
- \usepackage[labelsep=period]{caption}
- \usepackage[labelfont=bf]{caption}
- \captionsetup{justification=raggedright,singlelinecheck=false}

@@

muni.latest <- read.csv("muni_2021.csv", sep = ";" , header = F)
colnames(muni.latest) <- c('Area', 'Prop.Urban.Areas', 'Pop', 'Pop.Change', 'Prop.Below15', 'Prop.15to64', 'Prop.Over64', 'Prop.Swedish', 'Prop.Foreign', 'Pop.Growth', 'Migr.Gain', 'Families', 'Households', 'Prop.Households.rowSmall', 'Prop.Households.Rent', 'Prop.Educ.Degree2', 'Prop.Educ.High', 'Employed', 'Empl.Rate', 'Prop.Households.Empl', 'Prop.Unempl', 'Prop.Pension', 'Support.Ratio', 'Jobs', 'Prop.Primary.Sector', 'Prop.Secondary.Sector', 'Prop.Services.Sector', 'Jobs.Self.Suff', 'Margin.Citizen', 'Loan.Citizen', 'Concern.Loan.Citizen', 'Educ.Cult.Citizen', 'Soc.Health.Citizen')

@@

muni.latest <- muni.latest %>% 
  arrange(desc(Pop))

@@

\tableofcontents

@@

```{r, echo=FALSE}

vars.orig <- c('Taajama-aste, %', 'Väkiluku', 'Väkiluvun muutos edellisestä vuodesta, %', 'Alle 15-vuotiaiden osuus väestöstä, %', '15-64 -vuotiaiden osuus väestöstä, %', 'Yli 64-vuotiaiden osuus väestöstä, %', 'Ruotsinkielisten osuus väestöstä, %', 'Ulkomaan kansalaisten osuus väestöstä, %', 'Syntyneiden enemmyys, henkilöä', 'Kuntien välinen muuttovoitto/-tappio, henkilöä', 'Perheiden lukumäärä', 'Asuntokuntien lukumäärä', 'Rivi- ja pientaloissa asuvien asuntokuntien osuus, %', 'Vuokra-asunnoissa asuvien asuntokuntien osuus, %', 'Vähintään toisen asteen tutkinnon suorittaneiden osuus 15 vuotta täyttäneistä, %', 'Korkea-asteen tutkinnon suorittaneiden osuus 15 vuotta täyttäneistä, %', 'Alueella asuvan työllisen työvoiman määrä', 'Työllisyysaste, %', 'Asuinkunnassaan työssäkäyvien osuus, %', 'Työttömien osuus työvoimasta, %', 'Eläkeläisten osuus väestöstä, %', 'Taloudellinen huoltosuhde', 'Alueella olevien työpaikkojen lukumäärä', 'Alkutuotannon työpaikkojen osuus, %', 'Jalostuksen työpaikkojen osuus, %', 'Palvelujen työpaikkojen osuus, %', 'Työpaikkaomavaraisuus', 'Vuosikate, euroa/asukas', 'Lainakanta, euroa/asukas', 'Konsernin lainakanta, euroa/asukas', 'Opetus- ja kulttuuritoiminta yhteensä, nettokäyttökustannukset, euroa/asukas', 'Sosiaali- ja terveystoiminta yhteensä, nettokäyttökustannukset, euroa/asukas')

vars <-  c('Prop.Urban.Areas', 'Pop', 'Pop.Change', 'Prop.Below15', 'Prop.15to64', 'Prop.Over64', 'Prop.Swedish', 'Prop.Foreign', 'Pop.Growth', 'Migr.Gain', 'Families', 'Households', 'Prop.Households.rowSmall', 'Prop.Households.Rent', 'Prop.Educ.Degree2', 'Prop.Educ.High', 'Employed', 'Empl.Rate', 'Prop.Households.Empl', 'Prop.Unempl', 'Prop.Pension', 'Support.Ratio', 'Jobs', 'Prop.Primary.Sector', 'Prop.Secondary.Sector', 'Prop.Services.Sector', 'Jobs.Self.Suff', 'Margin.Citizen', 'Loan.Citizen', 'Concern.Loan.Citizen', 'Educ.Cult.Citizen', 'Soc.Health.Citizen')

explanations <- rep("NA", length(vars))

var.df <- data.frame(vars.orig, vars, explanations)

kable(var.df, format = "latex", align = "lll", col.names = c("Original", "Modified", "Explanation"), caption = "Variable Descriptions") %>%
  row_spec(0, bold = T) %>% 
  kable_styling(latex_options = c("HOLD_position", "scale_down"), position = "left")

@@

```{r, echo=FALSE}
stats <- muni.latest[, -1] %>% apply(2, summary) %>% t() %>% as.data.frame()
stats$Std <- muni.latest[, -1] %>% apply(2, sd)

@@

kable(stats.df, format = "latex", align = "lrrrrrr", caption = "Descriptive Statistics", digits = 1, col.names = c("Var", "Min", "1st Qu.", "Median", "Mean", "3rd Qu.", "Max", "Std.")) %>% 
  row_spec(0, bold = T) %>% 
  kable_styling(latex_options = "HOLD_position", position = "left")

@@

ages <- muni.latest %>% 
  select(Prop.Below15, Prop.15to64, Prop.Over64)
ages <- melt(ages)

living <- muni.latest %>% 
  select(Prop.Households.rowSmall, Prop.Households.Rent)
living <- melt(living)

employment <- muni.latest %>% 
  select(Empl.Rate, Prop.Unempl, Prop.Pension)
employment <- melt(employment)

@@

```{r, echo=FALSE, fig.cap='Probability density estimates for age groups', position, out.width='80%'}

ggplot(ages, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_minimal() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Age Groups") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r, echo=FALSE, fig.cap='Probability density estimates for living types', out.width='80%'}

ggplot(living, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_minimal() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Living Types") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r, echo=FALSE, fig.cap='Probability density estimates for employment groups', out.width='80%'}

ggplot(employment, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_minimal() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Employment Groups") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r, echo=FALSE, fig.cap='Probability density estimates for economic sectors', out.width='80%'}

ggplot(living, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_minimal() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Economic Sectors") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r, echo=FALSE, fig.dim=c(18,18), fig.cap="Heatmap of Pearson'n correlation between all variables", optipng = '-o7'}
corr <- cor(muni.latest[, -1])
ggcorrplot(corr, hc.order = TRUE, type = "full", lab = TRUE, title = "Heatmap of Correlations", lab_size = 3.5)

@@

@@

```{r, echo=FALSE, fig.cap="Variance explained by each Princial Component (PC) up to the 20th PC. This type of a plot is also known as a Scree Plot", optipng = '-o7'}

@@

ggplot(scree.df) +
  geom_bar(aes(x = PC, y = Prop.var, fill = Included), stat = "identity") +
  scale_fill_manual(values = c("#440154", "#3dbc74")) +
  geom_line(aes(x = PC, y = Cum.prop)) +
  geom_point(aes(x = PC, y = Cum.prop)) +
  geom_hline(yintercept = 0.75, linetype = 2) +
  annotate("text", x=10.5, y=0.78, label= "75% of variance explained", size = 3) +
  labs(x = "Principal Component", y = "Variance Explained (%)", title = "PCA Scree Plot") +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_minimal()

@@

```{r, echo=FALSE, message=FALSE, fig.cap="Score plot for all municipalities that have a top 5 absolute loading in any of PC1 and PC2", fig.align='center'}

@@

ggplot(loads.reduced, aes(x = PC1, y = PC2, label = Area)) +
  geom_point() +
  geom_text_repel(size = 3) +
  xlim(c(-25,10)) +
  ylim(c(-20,15)) +
  theme_minimal() +
  ggtitle("PCA Scores")

@@

```{r, warning=FALSE, echo=FALSE, fig.cap="Loading directions for PC1 and PC2", fig.align='center'}

ggplot2::autoplot(muni.pca, 
                  alpha = 0,
                  loadings = T, 
                  loadings.colour = "black", 
                  loadings.label = T, 
                  loadings.label.colour = "red",
                  loadings.label.size = 2.5, 
                  loadings.label.repel = T) +
  scale_x_continuous(limits = c(-0.4, 0.4)) +
  scale_y_continuous(limits = c(-0.4, 0.4)) +
  theme_minimal() +
  labs(x = "Standardized PC1 (33.6% explained var.)",
       y = "Standardized PC1 (20.9% explained var.)",
       title = "PCA Loadings")

@@

```{r, echo=FALSE, message=FALSE, fig.cap="Score plot for all municipalities that have a top 5 absolute loading in any of PC3 and PC4", fig.align='center'}

@@

ggplot(loads.reduced, aes(x = PC3, y = PC4, label = Area)) +
  geom_point() +
  geom_text_repel(size = 3) +
  xlim(c(-5,5)) +
  ylim(c(-8,8)) +
  theme_minimal() +
  ggtitle("PCA Scores")

@@

```{r, warning=FALSE, echo=FALSE, fig.cap="Loading directions for PC3 and PC4", fig.align='center'}

ggplot2::autoplot(muni.pca,
                  x = 3,
                  y = 4,
                  alpha = 0,
                  loadings = T, 
                  loadings.colour = "black", 
                  loadings.label = T, 
                  loadings.label.colour = "red",
                  loadings.label.size = 2.5, 
                  loadings.label.repel = T) +
  scale_x_continuous(limits = c(-0.3, 0.3)) +
  scale_y_continuous(limits = c(-0.3, 0.3)) +
  theme_minimal() +
  labs(x = "Standardized PC3 (10.2% explained var.)",
       y = "Standardized PC4 (7.2% explained var.)",
       title = "PCA Loadings")

>>>>

---
title: "PCA on Finnish Municipalities"
author: "Niko Miller"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: 
  bookdown::pdf_document2:
    keep_tex: false
    number_sections: true
    fig_caption: true
    citation_package: natbib
bibliography: references.bib
header-includes:
- \pagenumbering{gobble}
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsfonts}
- \usepackage{mathtools}
- \usepackage{graphicx}
- \usepackage{placeins}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{framed}
- \usepackage[font={small}]{caption}
- \usepackage[font={small,it}]{subcaption}
- \usepackage{hyperref}
- \usepackage[labelsep=period]{caption}
- \usepackage[labelfont=bf]{caption}
- \captionsetup{justification=raggedright,singlelinecheck=false}

@@

muni.latest <- read.csv("muni_2021.csv", sep = ";" , header = F)
colnames(muni.latest) <- c('Area',
                           'Deg.Urbanisation',
                           'Popul',
                           'Popul.Growth',
                           'Prop.Under15',
                           'Prop.15to64',
                           'Prop.Over64',
                           'Prop.Swedish',
                           'Prop.Foreign',
                           'Excess.Births',
                           'Migr.Gain',
                           'Families',
                           'Households',
                           'Prop.Households.Terr.Det',
                           'Prop.Households.Rental',
                           'Prop.Educ.Degree2',
                           'Prop.Educ.Degree3',
                           'Labour.Force',
                           'Empl.Rate',
                           'Prop.Empl.Muni',
                           'Prop.Unempl',
                           'Prop.Pensioners',
                           'Depend.Ratio',
                           'Jobs.Muni',
                           'Prop.Primary.Sector',
                           'Prop.Secondary.Sector',
                           'Prop.Services.Sector',
                           'Jobs.Self.Suff',
                           'Contr.Margin',
                           'Loan.Stock',
                           'Group.Loan.Stock',
                           'Educ.Cult.Activity',
                           'Soc.Health.Activity')

@@

muni.latest <- muni.latest %>% 
  arrange(desc(Popul))

@@

<!-- \tableofcontents -->

@@

Table \@ref(tab:vardesc) describes the variables used in this study.

```{r vardesc, echo=FALSE}

vars.orig <- c('Degree of urbanisation',
               'Population',
               'Population change from the previous year',
               'Share of persons aged under 15 of the population',
               'Share of persons aged 15 to 64 of the population',
               'Share of persons aged over 64 of the population',
               'Share of Swedish-speakers of the population',
               'Share of foreign citizens of the population',
               'Excess of births',
               'Intermunicipial migration gain/loss',
               'Number of families',
               'Number of household-dwelling units',
               'Share of household-dwelling units living in terraced houses and detached houses',
               'Share of household-dwelling units living in rental dwellings',
               'Share of persons aged 15 or over with at least upper secondary qualifications',
               'Share of persons aged 15 or over with tertiary level qualifications',
               'Employed labour force resident in the area',
               'Employment rate',
               'Share of persons working in their municipality of residence',
               'Proportion of unemployed among the labour force',
               'Proportion of pensioners of the population',
               'Economic dependency ratio',
               'Number of workplaces in the area',
               'Share of workplaces in primary production',
               'Share of workplaces in secondary production',
               'Share of workplaces in services',
               'Workplace self-sufficiency',
               'Annual contribution margin',
               'Loan stock',
               'Group loan stock',
               'Educational and cultural activities',
               'Social and health care activities')

vars <-  c('Deg.Urbanisation',
           'Popul',
           'Popul.Growth',
           'Prop.Under15',
           'Prop.15to64',
           'Prop.Over64',
           'Prop.Swedish',
           'Prop.Foreign',
           'Excess.Births',
           'Migr.Gain',
           'Families',
           'Households',
           'Prop.Households.Terr.Det',
           'Prop.Households.Rental',
           'Prop.Educ.Degree2',
           'Prop.Educ.Degree3',
           'Labour.Force',
           'Empl.Rate',
           'Prop.Empl.Muni',
           'Prop.Unempl',
           'Prop.Pensioners',
           'Depend.Ratio',
           'Jobs.Muni',
           'Prop.Primary.Sector',
           'Prop.Secondary.Sector',
           'Prop.Services.Sector',
           'Jobs.Self.Suff',
           'Contr.Margin',
           'Loan.Stock',
           'Group.Loan.Stock',
           'Educ.Cult.Activity',
           'Soc.Health.Activity')

years <- c(2020,
           2021,
           2021,
           2021,
           2021,
           2021,
           2021,
           2021,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2020,
           2019,
           2020,
           2020,
           2020,
           2019,
           2019,
           2019,
           2019,
           2019,
           2020,
           2020,
           2020,
           2020,
           2020)

units <- c("%",
           "Number",
           "%",
           "%",
           "%",
           "%",
           "%",
           "%",
           "Persons",
           "Persons",
           "Number",
           "Number",
           "%",
           "%",
           "%",
           "%",
           "Number",
           "%",
           "%",
           "%",
           "%",
           "Ratio",
           "Number",
           "%",
           "%",
           "%",
           "Ratio",
           "EUR per capita",
           "EUR per capita",
           "EUR per capita",
           "EUR per capita",
           "EUR per capita")

var.df <- data.frame(vars, vars.orig, years, units)

kable(var.df, format = "latex", align = "lll", col.names = c("Variable", "Original Variable","Year","Unit"), caption = "Description of all used variables. Variable is the name of the variable that is used in this report. Original Variable is the original variable given in Tilastokeskus' dataset, Year is the timestamp of the data point, and Unit is the unit in which the variable is measured in.") %>%
  row_spec(0, bold = T) %>% 
  kable_styling(latex_options = c("HOLD_position", "scale_down"), position = "left")

@@

Table \@ref(tab:descstats) shows descriptive statistics for each variable.

```{r descstats, echo=FALSE}
stats <- muni.latest[, -1] %>% apply(2, summary) %>% t() %>% as.data.frame()
stats$Std <- muni.latest[, -1] %>% apply(2, sd)

@@

kable(stats.df, format = "latex", align = "lrrrrrr", caption = "Descriptive Statistics of the variables. Variable is the name of the variable, Min is the minimum, 1st Qu. is the 1st quartile, Median is the median, Mean is the arithmetic mean, 3rd Qu. is the 3rd quartile, Max is the maximum and Std. is the sample standard deviation.", digits = 1, col.names = c("Variable", "Min", "1st Qu.", "Median", "Mean", "3rd Qu.", "Max", "Std.")) %>% 
  row_spec(0, bold = T) %>% 
  kable_styling(latex_options = "HOLD_position", position = "left")

@@

ages <- muni.latest %>% 
  select(Prop.Under15, Prop.15to64, Prop.Over64)
ages <- melt(ages)

living <- muni.latest %>% 
  select(Prop.Households.Terr.Det, Prop.Households.Rental)
living <- melt(living)

employment <- muni.latest %>% 
  select(Empl.Rate, Prop.Unempl, Prop.Pensioners)
employment <- melt(employment)

@@

To analyze how variables are distributed, we use Kernel density estimation to estimate the probability densities for several variable groups.

Figure \@ref(fig:agekde) shows the probability density estimates for age groups. Age groups are shares of population aged below 15, 15 to 64, and over 64. We can see that ...

Figure \@ref(fig:housingkde) shows the probability density estimates for housing types. Housing types are shares of population living in either terraced or detached houses and in rental apartments. The Figure revels that ...

Figure \@ref(fig:employmentkde) shows the probability density estimates for employment measures. Employment measures are the employment rate, the proportion of population that are unemployed, and the proportion of population that are pensioners. The distributions show that ...

Figure \@ref(fig:sectorkde) shows the probability density estimates for the weights of economic sectors in the municipalities. Sectors are the primary-, secondary-, and services sectors. We can interpret that ...

```{r agekde, echo=FALSE, fig.cap='Probability density estimates for age groups', fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(ages, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Age Groups") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r housingkde, echo=FALSE, fig.cap='Probability density estimates for housing types', fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(living, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Housing Types") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r employmentkde, echo=FALSE, fig.cap='Probability density estimates for employment related measures',fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(employment, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Employment Measures") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r sectorkde, echo=FALSE, fig.cap='Probability density estimates for economic sectors', fig.align='left', out.width='80%', optipng = '-o7'}

ggplot(sectors, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_light() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Economic Sectors") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

To analyze bivariate dependencies in the data, we assess linear relationships between the variables with Pearson's correlation analysis. Figure \@ref(fig:corrplot) shows a heatmap of correlations between all variables.

The heatmap shows that ...

```{r corrplot, echo=FALSE, fig.dim=c(18,18), fig.cap="Heatmap of Pearson'n correlation between all variables", optipng = '-o7'}
corr <- cor(muni.latest[, -1])
ggcorrplot(corr, hc.order = TRUE, type = "full", lab = TRUE, title = "Heatmap of Correlations", lab_size = 3.5)

@@

To assess how much proportion of variance can be explained with the first few PCs, we create a scree plot. Figure \@ref(fig:screeplot) shows the resulting plot. We can see that we can explain close to 55% of all variation using only the first two PCs. Furthermore, close to 70% of the variation can be explained using the first four PCs.

@@

```{r screeplot, echo=FALSE, fig.cap="Variance explained by each Princial Component (PC) up to the 20th PC. This type of a plot is also known as a Scree Plot", optipng = '-o7'}

@@

ggplot(scree.df) +
  geom_bar(aes(x = PC, y = Prop.var, fill = Included), stat = "identity") +
  scale_fill_manual(values = c("#440154", "#3dbc74")) +
  geom_line(aes(x = PC, y = Cum.prop)) +
  geom_point(aes(x = PC, y = Cum.prop)) +
  geom_hline(yintercept = 0.75, linetype = 2) +
  annotate("text", x=10.5, y=0.78, label= "75% of variance explained", size = 3) +
  labs(x = "Principal Component", y = "Variance Explained (%)", title = "PCA Scree Plot") +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_light()

@@

Figure \@ref(fig:scoreplot1) shows the scores for PCs 1 and 2.

Figure \@ref(fig:loadingsplot1) shows the loading directions for PCs 1 and 2.

Figure \@ref(fig:scoreplot2) shows the scores for PCs 3 and 4.

Figure \@ref(fig:loadingsplot2) shows the loading directions for PCs 3 and 4.
```{r scoreplot1, echo=FALSE, message=FALSE, fig.cap="Score plot for all municipalities that have a top 5 absolute loading in any of PC1 and PC2", fig.align='center'}

@@

ggplot(loads.reduced, aes(x = PC1, y = PC2, label = Area)) +
  geom_point() +
  geom_text_repel(size = 3) +
  xlim(c(-25,10)) +
  ylim(c(-20,15)) +
  ggtitle("PCA Scores") +
  theme_light()

@@

```{r loadingsplot1, warning=FALSE, echo=FALSE, fig.cap="Loading directions for PC1 and PC2", fig.align='center'}

ggplot2::autoplot(muni.pca, 
                  alpha = 0,
                  loadings = T, 
                  loadings.colour = "black", 
                  loadings.label = T, 
                  loadings.label.colour = "red",
                  loadings.label.size = 2.5, 
                  loadings.label.repel = T) +
  scale_x_continuous(limits = c(-0.4, 0.4)) +
  scale_y_continuous(limits = c(-0.4, 0.4)) +
  labs(x = "Standardized PC1 (33.6% explained var.)",
       y = "Standardized PC1 (20.9% explained var.)",
       title = "PCA Loadings") +
  theme_light()

@@

```{r scoreplot2, echo=FALSE, message=FALSE, fig.cap="Score plot for all municipalities that have a top 5 absolute loading in any of PC3 and PC4", fig.align='center'}

@@

ggplot(loads.reduced, aes(x = PC3, y = PC4, label = Area)) +
  geom_point() +
  geom_text_repel(size = 3) +
  xlim(c(-5,5)) +
  ylim(c(-8,8)) +
  ggtitle("PCA Scores") +
  theme_light()

@@

```{r loadingsplot2, warning=FALSE, echo=FALSE, fig.cap="Loading directions for PC3 and PC4", fig.align='center'}

ggplot2::autoplot(muni.pca,
                  x = 3,
                  y = 4,
                  alpha = 0,
                  loadings = T, 
                  loadings.colour = "black", 
                  loadings.label = T, 
                  loadings.label.colour = "red",
                  loadings.label.size = 2.5, 
                  loadings.label.repel = T) +
  scale_x_continuous(limits = c(-0.3, 0.3)) +
  scale_y_continuous(limits = c(-0.3, 0.3)) +
  labs(x = "Standardized PC3 (10.2% explained var.)",
       y = "Standardized PC4 (7.2% explained var.)",
       title = "PCA Loadings") +
  theme_light()
